---
title: "More Flexible Model Run"
author: "H. Juliette T. Unwin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

Within this package there are 2 ways to run the deterministic base model. This vignette focuses on the 
`create_r_model` function, which gives the user maximum control of the input parameters. The second,  
`run_model` function,  solves the base model. For more infomation about the `run_model` function see this 
[vignette](https://mrc-ide.github.io/hanojoel/articles/run_model.html).

## Default use of function
The `create_r_model` function creates the odin deterministic malaria model in R.  It has the 
following default arguments:

1. Odin model path - `odin_model_path = system.file("extdata/odin_model.R", package = "hanojoel"`
1. The number of heterogeneity brakets - `het_brackets = 5`
2. The vector of age compartments - `age = c(0,0.25,0.5,0.75,1,1.25,1.5,1.75,2,3.5,5,7.5,10,15,20,30,40,50,60,70,80)`
3. The initial EIR - `init_EIR = 10`
4. The percentage of the population that gets treated - `init_ft = 0.4`
5. The country - `country = NULL`
6. The admin 2 unit - `admin2 = NULL`

The simplest example of using the function to plot the malaria prevelance for a country
with no seasonablity (found by setting both the country and admin2 to NULL) is:

```{r}
library(hanojoel)

wh <- hanojoel:::create_r_model()

# generates model functions with initial state data
mod <- wh$generator(user= wh$state, use_dde = TRUE)

# Runs the model
mod_run <- mod$run(t = 1:365)
out <- mod$transform_variables(mod_run)
plot(out$t,out$prev, main= "Prevalance", ylim = c(0.22, 0.24), type='l')
```

## Providing alternative arguments
It is posisble to change the default arguments for the `create_r_model` function as shown below: 
```{r, message = FALSE}
# create a vector of age categories
init_age <- c(0,0.25,0.5,0.75,1,1.25,1.5,1.75,2,3.5,5,7.5,10,15,20,30,40,50,60,70,80)

# provide a value of the annual EIR for this model run
init_EIR <- 10

# provide a string for the admin 2 unit that you want to use the seasonality profile for
admin_str <- "Tororo"

# provide the length of time (in days) that you want to run the model for
time_period <- 365*1

# provide a value for the proportion of cases that are treated (referred to as ft in the paper)
prop_treated <- 0.4

# creates the odin model using the default odin model file
wh <- hanojoel:::create_r_model(odin_model_path = system.file("extdata/odin_model.R", package = "hanojoel"),
                                het_brackets = 5,
                                age = init_age,
                                init_EIR = init_EIR,
                                init_ft = prop_treated,
                                country = "Uganda",
                                admin2 = "Tororo")

# generates model functions with initial state data
mod <- wh$generator(user= wh$state, use_dde = TRUE)

# Runs the model
mod_run <- mod$run(t = 1:365)
out <- mod$transform_variables(mod_run)
plot(out$t,out$prev, main="Prevalence", type='l')
```

It is also possible to change the odin model by editing the `odin_model_path`.  If the model is 
built within the package, such as `odin_model_hrp2.R`, change the path to:
```{r}
odin_model_path = system.file("extdata/odin_model_hrp2.R", package = "hanojoel")
```
If the odin model is only held locally, change the model path to:
```{r}
odin_model_path = "extdata/odin_model_hrp2.R"
```


## Adding additional optional arguments
It may be necessary to provide extra optional arguments to the odin model. For example if you
run the `odin_model_hrp2.R` model without any extra arguments it produces an error:
```{r, error=TRUE, purl = FALSE}
# creates the odin model using the default odin model file
wh <- hanojoel:::create_r_model(odin_model_path = system.file("extdata/odin_model_hrp2.R", 
                                                              package = "hanojoel"))
mod <- wh$generator(user= wh$state, use_dde = TRUE)
```

This is because the required user argument `hrp2_prop` has not been provided. This is done as follows:
```{r}
# creates the odin model using the default odin model file
wh <- hanojoel:::create_r_model(odin_model_path = system.file("extdata/odin_model_hrp2.R", 
                                                              package = "hanojoel"), 
                                hrp2_prop = 0.5)
mod <- wh$generator(user= wh$state, use_dde = TRUE)
```


## Changing seasonality parameters
It is possible to change the seasonality parameters by specifying a named admin unit or both
an admin unit and country to the `hanojoel:::create_r_model()` function.  The latter is better when the admin 
name is common. Fuzzy matching
is used to match the provided arguments with the data set provided in the package.  The data set
can be loaded into your workspace as follows:
```{r}
seasonality_data <- load_file("admin_units_seasonal.rds")
```


## Adding interventions
It is possible to vary the number of interventions considered in the model.  

No interventions are included in the default settings of the model so `num_int = 1.`  

Insecticide treated nets can be added to the default model as follows. Two intervention compartments are
necessary. Net usage is set to 50% in this example and is turned on
365 days into the simulation:
```{r}
wh <- hanojoel:::create_r_model(odin_model_path = system.file("extdata/odin_model.R", package = "hanojoel"),
                                het_brackets = 5,
                                age = init_age,
                                init_EIR = init_EIR,
                                init_ft = prop_treated,
                                country = "Uganda",
                                admin2 = "Tororo",
                                num_int = 2,
                                ITN_IRS_on = 365,
                                itn_cov = 0.5)

# generates model functions with initial state data
mod <- wh$generator(user= wh$state, use_dde = TRUE)

# Runs the model
mod_run <- mod$run(t = 1:700)
out <- mod$transform_variables(mod_run)
plot(out$t,out$prev, main="Prevalence", type='l')
```

Indoor residual spraying (IRS) can be added to default model as follows. It requires four intervention 
compartments. Net usage is set to 50% in this example, IRS usage is 25% and both are turned on
365 days into the simulation:


```{r, message = FALSE}
wh <- hanojoel:::create_r_model(odin_model_path = system.file("extdata/odin_model.R", package = "hanojoel"),
                                het_brackets = 5,
                                age = init_age,
                                init_EIR = init_EIR,
                                init_ft = prop_treated,
                                country = "Uganda",
                                admin2 = "Tororo",
                                num_int = 4,
                                ITN_IRS_on = 365,
                                itn_cov = 0.5,
                                irs_cov = 0.25)

# generates model functions with initial state data
mod <- wh$generator(user= wh$state, use_dde = TRUE)

# Runs the model
mod_run <- mod$run(t = 1:700)
out <- mod$transform_variables(mod_run)
plot(out$t,out$prev, main="Prevalence", type='l')
```

Other interventions can be added in new model files specified with the `odin_model_path` variable.  Any
changes to the equilibrium solution should be made between the `hanojoel:::create_r_model()` and
`wh$generator()` function calls and not through edits to the equilibrium solution in the model.
