---
title: "More Flexible Model Run"
author: "H. Juliette T. Unwin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

Within this package there are 2 ways to run the deterministic base model. This vignette focuses on the 
`create_r_model` function, which gives the user maximum control of the input parameters. The second,  
`run_model` function,  solves the base model. For more infomation about the `run_model` function see this 
[vignette](https://mrc-ide.github.io/hanojoel/articles/run_model.html).

## Default use of function
The `create_r_model` function creates the odin deterministic malaria model in R.  It has the 
following default arguments:

1. Odin model path - `odin_model_path = system.file("extdata/odin_model.R", package = "hanojoel"`
1. The number of heterogeneity brakets - `het_brackets = 5`
2. The vector of age compartments - `age = c(0,0.25,0.5,0.75,1,1.25,1.5,1.75,2,3.5,5,7.5,10,15,20,30,40,50,60,70,80)`
3. The initial EIR - `init_EIR = 10`
4. The percentage of the population that gets treated - `init_ft = 0.4`
5. The country - `country = NULL`
6. The admin 2 unit - `admin2 = NULL`

The simplest example of using the function to plot the malaria prevelance for a country
with no seasonablity (found by setting both the country and admin2 to NULL) is:

```{r}
library(hanojoel)

wh <- hanojoel:::create_r_model()

# generates model functions with initial state data
mod <- wh$generator(user= wh$state, use_dde = TRUE)

# Runs the model
mod_run <- mod$run(t = 1:365)
out <- mod$transform_variables(mod_run)
plot(out$t,out$prev, main= "Prevalance")
```

## Providing alternative arguments
It is posisble to change the default arguments for the `create_r_model` function as shown below: 
```{r}
# create a vector of age categories
init_age <- c(0,0.25,0.5,0.75,1,1.25,1.5,1.75,2,3.5,5,7.5,10,15,20,30,40,50,60,70,80)

# provide a value of the annual EIR for this model run
init_EIR <- 10

# provide a string for the admin 2 unit that you want to use the seasonality profile for
admin_str <- "Tororo"

# provide the length of time (in days) that you want to run the model for
time_period <- 365*1

# provide a value for the proportion of cases that are treated (referred to as ft in the paper)
prop_treated <- 0.4

# creates the odin model using the default odin model file
wh <- hanojoel:::create_r_model(odin_model_path = system.file("extdata/odin_model.R", package = "hanojoel"),
                                het_brackets = 5,
                                age = init_age,
                                init_EIR = init_EIR,
                                init_ft = prop_treated,
                                country = "Uganda",
                                admin2 = "Tororo")

# generates model functions with initial state data
mod <- wh$generator(user= wh$state, use_dde = TRUE)

# Runs the model
mod_run <- mod$run(t = 1:365)
out <- mod$transform_variables(mod_run)
plot(out$t,out$prev, main="Prevalance")
```

It is also possible to change the odin model by editing the `odin_model_path`.  If the model is 
built within the package, such as `odin_model_hrp2.R`, change the path to:
```{r}
odin_model_path = system.file("extdata/odin_model_hrp2.R", package = "hanojoel")
```
If the odin model is only held locally, change the model path to:
```{r}
odin_model_path = "extdata/odin_model_hrp2.R"
```

## Adding additional optional arguments
It may be necessary to provide extra optional arguments to the odin model. For example if you
run the `odin_model_hrp2.R` model without any extra arguments it produces an error:
```{r, error=TRUE, purl = FALSE}
# creates the odin model using the default odin model file
wh <- hanojoel:::create_r_model(odin_model_path = system.file("extdata/odin_model_hrp2.R", 
                                                              package = "hanojoel"))
mod <- wh$generator(user= wh$state, use_dde = TRUE)
```

This is because the required user argument `hrp2_prop` has not been provided. This is done as follows:
```{r}
# creates the odin model using the default odin model file
wh <- hanojoel:::create_r_model(odin_model_path = system.file("extdata/odin_model_hrp2.R", 
                                                              package = "hanojoel"), 
                                hrp2_prop = 0.5)
mod <- wh$generator(user= wh$state, use_dde = TRUE)
```
