---
title: "Run Base Model"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

Within this package there are 2 ways to run the deterministic model:

1. `run_modelb`
2. `create_r_model`

`run_model` will run the base model, which can be found in 
`inst/extdata/odin_model.R`. `run_model` is designed to quickly simulate 
transmission dynamics, and as such there are fewer parameters that you are able 
to change. Whereas `create_r_model` allows you to specify
what model you would like to run. For more information please see the other 
[vignette](https://mrc-ide.github.io/hanojoel/articles/anemia.html)

## Default Use

`run_model` can be used to simulate the base model and will return a list of 
2 elements.

1. `dat` - the simulation output
2. `plot` - a plot of the 6 human compartments

Without specifying any function arguments, `create_r_model` will run a 
simulation for 1 year at an EIR of 10, with a treatment coverage equal to 40% 
for Tororo, Uganda. 

```{r, fig.show='hold'}
library(hanojoel)
out <- run_model()
names(out)
out$plot

```

## Passing Variables

To change simulation settings we pass in one of 5 function arguments:

1. age Vector of age brackets.
2. EIR Numeric of annual EIR. 
3. ft Numeric of proportion of symptomatic cases recieving treatment.
4. time Numeric of length of time that the model will run for in days.
5. admin2 Character of admin unit. 

For example, to simulate Thies, Senegal at an EIR of 1 for 2 years at a 
treatment coverage of 60%:

```{r, echo=FALSE, results='asis'}
thies <- run_model(EIR = 1, ft = 0.6, admin2 = "Thies", time = 2*365)
thies$plot
```

We can also then plot other model outputs by interacting with the `dat` 
data frame returned by `run_model`:

```{r, echo=FALSE, results='asis'}
plot(thies$dat$Sv, ylab = "Susceptible Adult Mosquitoes", xlab = "Time (days)")
```
