---
title: "Simplest Model Run"
author: "OJ Watson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

Within this package there are 2 ways to run the deterministic base model:

1. `run_model`
2. `create_r_model`

The `run_model` function solves the base model, which can be found in 
`inst/extdata/odin_model.R`. This function is designed to quickly simulate 
transmission dynamics, and as such there are fewer parameters that you are able 
to change. Whereas, `create_r_model` allows you to specify
what model you would like to run. For more information please see the other 
[vignette](https://mrc-ide.github.io/hanojoel/articles/anemia.html).

## Default Use
This vignette focuses on using the `run_model` function to simulate the base 
model. It returns the following list with 2 elements:

1. `dat` - the simulation output.
2. `plot` - a plot of the 6 human compartments.

As illustrated below, without specifying any function arguments, `create_r_model` 
will run a simulation for 1 year at an EIR of 10, with a treatment coverage equal to 40% 
for Tororo, Uganda.

```{r, fig.show='hold', message = FALSE}
library(hanojoel)
out <- run_model()
names(out)
out$plot

```

## Passing Variables

To change simulation settings there are 5 optional function arguments:

1. `age` - Vector of age brackets.
2. `EIR` -  Numeric of annual EIR. 
3. `ft` - Numeric of proportion of symptomatic cases recieving treatment.
4. `admin2` - Character of admin unit. 
5. `time` - Numeric of length of time that the model will run for in days.

For example, to simulate Thies, Senegal for a given vector of ages, at an EIR of 1, for 2 years with 
60% treatment coverage:

```{r, echo=FALSE, results='asis', message = FALSE}
thies <- run_model(age = c(0, 1, 2, 3, 4, 5, 10, 20, 40, 60), 
                   EIR = 1, 
                   ft = 0.6, 
                   admin2 = "Thies", 
                   time = 2*365)
thies$plot
```

Other model outputs, such as the number of suceptible adult mosquitoes, can be plotted by interacting with the `dat` data frame returned by `run_model`:

```{r, echo=FALSE, results='asis', message = FALSE}
plot(thies$dat$Sv, ylab = "Susceptible Adult Mosquitoes", xlab = "Time (days)", main="Number of susceptible mosquitoes")
```
