---
title: "Writing your own model"
author: "Joel Hellewell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Writing your own model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

You might want to write your own extension to the malaria models, in this vignette I will try and cover the easiest way of doing this.

You should start by cloning this github repository to your own computer. If you are new to R package development the next best step is to open the `ICDMM.Rproj` file using RStudio. 

## Adding your own odin model

If you write some new code in a file called `example_model.R`, you should save it in `inst/odin` with the other models. You then need to run `odin::odin_package()`, with the directory path of the project inside the parenthesis, this will add a compiled version of your odin model to the package files. 

## Other code changes

You will need to add your model name to the following bit of code in line 44 of `R/run_model.R`:
```{r, fig.show='hold', message = FALSE, eval= FALSE}
  generator <- switch(model,
    "odin_model" = odin_model,
    "odin_model_emanators" = odin_model_emanators,
    "odin_model_hrp2" = odin_model_hrp2,
    "odin_model_IVM_SMChet" = odin_model_IVM_SMChet,
    "example" = example,
    stop(sprintf("Unknown model '%s'", model)))

```
This will allow for you to eventually run `run_model(model = "example")` to run your odin model. 

## Rebuild package
The final step is to rebuild the R package, you can do this by pressing "Install and Restart" in the "Build" panel in RStudio (usually in the top right panels) or the usual keyboard is ctrl + shift + b (or cmd + shift + b on a mac)

## Working with github
The easiest way to develop your new package using github will be to make a new branch. You can do this using the "New Branch" button in the RStudio interface or by running `git checkout -b <branch name>`. You can then push this new branch to github to do version control but your changes won't affect anything on the main branch. When you have finished working on your model you can use github to create a pull request and we can merge it with the main branch so that everyone downloading the package also recieves your model as part of it.

